---
title: "College Football Recruiting"
author: "Trent Bellinger"
date: "2023-03-21"
output: html_document
---

``` {r}
library(devtools)
library(ggplot2)
library(lmtest)
library(sandwich)
library(car)
library(AER)
library(broom)
library(leaps)
library(Boruta)
library("PerformanceAnalytics")
library(corrgram)
library(lmvar)

transfer <- read.csv("transfers.csv", header = TRUE)
recruits <- read.csv("CFB_Recruits_Data.csv", header = TRUE)
data <- read.csv("cfb2022.csv", header = TRUE)

transfer2021 <- transfer[transfer$season == 2021, ]
transfer2022 <- transfer[transfer$season == 2022, ]
transfer2023 <- transfer[transfer$season == 2023, ]
```

Creating a new dataset for each team's transfers for each year.
``` {r}
num2021 <- c()
trans2021 <- tapply(transfer2021$destination, transfer2021$destination, length)
for (i in seq_along(data$Team)) {
  if (any(names(trans2021) == data$Team[i])) {
    index <- which(names(trans2021) == data$Team[i])
    num2021[i] <- trans2021[index]
  } else {
    num2021[i] <- 0
  }
}
num2021
trans2021

num2022 <- c()
trans2022 <- tapply(transfer2022$destination, transfer2022$destination, length)
for (i in seq_along(data$Team)) {
  if (any(names(trans2022) == data$Team[i])) {
    index <- which(names(trans2022) == data$Team[i])
    num2022[i] <- trans2022[index]
  } else {
    num2022[i] <- 0
  }
}
num2022
trans2022

num2023 <- c()
trans2023 <- tapply(transfer2023$destination, transfer2023$destination, length)
for (i in seq_along(data$Team)) {
  if (any(names(trans2023) == data$Team[i])) {
    index <- which(names(trans2023) == data$Team[i])
    num2023[i] <- trans2023[index]
  } else {
    num2023[i] <- 0
  }
}
num2023
trans2023

stars2021 <- c()
trans2021_stars <- tapply(transfer2021$stars, transfer2021$destination, mean)
for (i in seq_along(data$Team)) {
  if (any(names(trans2021_stars) == data$Team[i])) {
    index <- which(names(trans2021_stars) == data$Team[i])
    stars2021[i] <- trans2021_stars[index]
  } else {
    stars2021[i] <- 0
  }
}
stars2021
trans2021_stars

stars2022 <- c()
trans2022_stars <- tapply(transfer2022$stars, transfer2022$destination, mean)
for (i in seq_along(data$Team)) {
  if (any(names(trans2022_stars) == data$Team[i])) {
    index <- which(names(trans2022_stars) == data$Team[i])
    stars2022[i] <- trans2022_stars[index]
  } else {
    stars2022[i] <- 0
  }
}
stars2022
trans2022_stars

stars2023 <- c()
trans2023_stars <- tapply(transfer2023$stars, transfer2023$destination, mean)
for (i in seq_along(data$Team)) {
  if (any(names(trans2023_stars) == data$Team[i])) {
    index <- which(names(trans2023_stars) == data$Team[i])
    stars2023[i] <- trans2023_stars[index]
  } else {
    stars2023[i] <- 0
  }
}
stars2023
trans2023_stars

team_transfers <- data.frame("Team" = data$Team, "Conference" = data$Conference, 
                             "Transfers2021" = num2021, 
                             "Transfers2022" = num2022, "Transfers2023" = num2023, 
                             "AvgStars2021" = stars2021, "AvgStars2022" = stars2022, 
                             "AvgStars2023" = stars2023, "Wins2022" = data$Win)
head(team_transfers)

write.csv(team_transfers, "/Users/trentbellinger/Desktop/College Football/team transfers.csv", 
          row.names = FALSE)
```

```{r}
team_transfers <- read.csv("team transfers.csv", header = TRUE)

# average stars of transfers before 2022 season vs. wins in 2022 season
plot(x = team_transfers$Wins2022, y = team_transfers$AvgStars2022, cex = 0.7, pch = 19)

# number of transfers before 2022 season vs. wins in 2022 season
plot(x = team_transfers$Wins2022, y = team_transfers$Transfers2022, cex = 0.7, pch = 19)
```

Now creating a new dataset for recruits using only FBS Schools

```{r}
head(recruits)

# filter out recruits who did not commit to a school
recruits_new <- recruits[!recruits$committed_to == "", ]

# filter out any recruits that did not commit to the FBS schools we have data for
schools <- names(table(recruits_new$committed_to))
for (i in seq_along(schools)) {
  if (!any(data$Team == schools[i])) {
    index <- which(recruits_new$committed_to == schools[i])
    recruits_new <- recruits_new[-index, ]
  }
}
head(recruits_new)

write.csv(recruits_new, "/Users/trentbellinger/Desktop/College Football/recruits.csv", 
          row.names = FALSE)
```

Create a New Dataset for Recruits on Every Team
```{r}
names(table(recruits$year))

recruits2012 <- recruits[recruits$year == 2012, ]
recruits2013 <- recruits[recruits$year == 2013, ]
recruits2014 <- recruits[recruits$year == 2014, ]
recruits2015 <- recruits[recruits$year == 2015, ]
recruits2016 <- recruits[recruits$year == 2016, ]
recruits2017 <- recruits[recruits$year == 2017, ]
recruits2018 <- recruits[recruits$year == 2018, ]
recruits2019 <- recruits[recruits$year == 2019, ]
recruits2020 <- recruits[recruits$year == 2020, ]
recruits2021 <- recruits[recruits$year == 2021, ]
recruits2022 <- recruits[recruits$year == 2022, ]
recruits2023 <- recruits[recruits$year == 2023, ]

num2023 <- c()
rec2023 <- tapply(recruits2023$committed_to, recruits2023$committed_to, length)
for (i in seq_along(data$Team)) {
  if (any(names(rec2023) == data$Team[i])) {
    index <- which(names(rec2023) == data$Team[i])
    num2023[i] <- rec2023[index]
  } else {
    num2023[i] <- 0
  }
}
num2023
rec2023

rating2022 <- c()
rec2022_rating <- tapply(recruits2022$rating, recruits2022$committed_to, mean)
for (i in seq_along(data$Team)) {
  if (any(names(rec2022_rating) == data$Team[i])) {
    index <- which(names(rec2022_rating) == data$Team[i])
    rating2022[i] <- rec2022_rating[index]
  } else {
    rating2022[i] <- 0
  }
}

stars2012 <- tapply(recruits2012$stars, recruits2012$committed_to, mean)
length(stars2012)
# 2013 done using above loop
stars2014 <- tapply(recruits2014$stars, recruits2014$committed_to, mean)
length(stars2014)
stars2015 <- tapply(recruits2015$stars, recruits2015$committed_to, mean)
length(stars2015)
stars2016 <- tapply(recruits2016$stars, recruits2016$committed_to, mean)
length(stars2016)
stars2017 <- tapply(recruits2017$stars, recruits2017$committed_to, mean)
length(stars2017)
stars2018 <- tapply(recruits2018$stars, recruits2018$committed_to, mean)
length(stars2018)
stars2019 <- tapply(recruits2019$stars, recruits2019$committed_to, mean)
length(stars2019)
stars2020 <- tapply(recruits2020$stars, recruits2020$committed_to, mean)
length(stars2020)
# 2021 and 2022 done using above loop
stars2023 <- tapply(recruits2023$stars, recruits2023$committed_to, mean)
length(stars2023)

rating2012 <- tapply(recruits2012$rating, recruits2012$committed_to, mean)
# 2013 done using above loop
rating2014 <- tapply(recruits2014$rating, recruits2014$committed_to, mean)
rating2015 <- tapply(recruits2015$rating, recruits2015$committed_to, mean)
rating2016 <- tapply(recruits2016$rating, recruits2016$committed_to, mean)
rating2017 <- tapply(recruits2017$rating, recruits2017$committed_to, mean)
rating2018 <- tapply(recruits2018$rating, recruits2018$committed_to, mean)
rating2019 <- tapply(recruits2019$rating, recruits2019$committed_to, mean)
rating2020 <- tapply(recruits2020$rating, recruits2020$committed_to, mean)
# 2021 and 2022 done using above loop
rating2023 <- tapply(recruits2023$rating, recruits2023$committed_to, mean)

team_recruits <- data.frame("Team" = data$Team, "Conference" = data$Conference, 
                            "Recruits2012" = num2012, 
                            "AvgStars2012" = stars2012, "AvgRating2012" = rating2012, 
                            "Recruits2013" = num2013, "AvgStars2013" = stars2013, 
                            "AvgRating2013" = rating2013, "Recruits2014" = num2014, 
                             "AvgStars2014" = stars2014, "AvgRating2014" = rating2014,
                            "Recruits2015" = num2015, "AvgStars2015" = stars2015, 
                            "AvgRating2015" = rating2015, "Recruits2016" = num2016, 
                             "AvgStars2016" = stars2016, "AvgRating2016" = rating2016,
                            "Recruits2017" = num2017, "AvgStars2017" = stars2017, 
                            "AvgRating2017" = rating2017, "Recruits2018" = num2018, 
                             "AvgStars2018" = stars2018, "AvgRating2018" = rating2018, 
                            "Recruits2019" = num2019, "AvgStars2012" = stars2019, 
                            "AvgRating2019" = rating2019, "Recruits2020" = num2020, 
                             "AvgStars2020" = stars2020, "AvgRating2020" = rating2020,
                            "Recruits2021" = num2021, "AvgStars2012" = stars2021, 
                            "AvgRating2021" = rating2021, "Recruits2022" = num2022, 
                             "AvgStars2022" = stars2022, "AvgRating2022" = rating2022,  
                            "Recruits2023" = num2023, "AvgStars2023" = stars2023, 
                            "AvgRating2023" = rating2023, "Wins2022" = data$Win)

write.csv(team_recruits, "/Users/trentbellinger/Desktop/College Football/team recruits.csv", 
          row.names = FALSE)
```

Pac-12 only data
```{r}
pac12 <- data[data$Conference == "Pac-12", ]

pac12_recruits <- data.frame()
for (i in seq_along(recruits_new$committed_to)) {
  if (any(pac12$Team == recruits_new$committed_to[i])) {
    pac12_recruits <- rbind(pac12_recruits, recruits_new[i, ])
  }
}

pac12_recruits2019 <- pac12_recruits[pac12_recruits$year == 2019, ]
plot(x = pac12$Win, y = tapply(pac12_recruits$rating, pac12_recruits$committed_to, mean), 
     pch = 19, cex = 0.7)
```

Power 5 Only Data
```{r}
pwr5 <- data[data$Conference == "Pac-12" | data$Conference == "ACC" | data$Conference == "SEC" | 
             data$Conference == "Big 12" | data$Conference == "Big Ten", ]

pwr5_recruits <- data.frame()
for (i in seq_along(recruits$committed_to)) {
  if (any(pwr5$Team == recruits$committed_to[i])) {
    pwr5_recruits <- rbind(pwr5_recruits, recruits[i, ])
  }
}

pwr5_recruits2019 <- pwr5_recruits[pwr5_recruits$year == 2019, ]
pwr5_recruits2020 <- pwr5_recruits[pwr5_recruits$year == 2020, ]
plot(x = pwr5$Win, y = tapply(pwr5_recruits2020$rating, pwr5_recruits2020$committed_to, mean), 
     pch = 19, cex = 0.7)
```

Looking to see which recruiting class most impacts team success
```{r}
team_recruits <- read.csv("team recruits.csv", header = TRUE)

colnames(team_recruits)
model1 <- lm(Wins2022 ~ Recruits2018 + AvgStars2018 + AvgRating2018 + Recruits2019 + AvgStars2019 +
            AvgRating2019 + Recruits2020 + AvgStars2020 + AvgRating2020 + Recruits2021 + 
            AvgStars2021 + AvgRating2021 + Recruits2022 + AvgStars2022 + AvgRating2022, 
            data = team_recruits)
summary(model) 
# looks like avegare stars for 2019 and number of recruits for 2019 are most significant

model2 <- lm(Wins2022 ~ AvgStars2019, data = team_recruits)
summary(model2)
plot(x = team_recruits$AvgStars2019, y = team_recruits$Wins2022, cex = 0.7, pch = 19)
```

App that allows users to see recruiting class rankings
```{r}
library(shiny)
recruits <- read.csv("team recruits.csv", header = TRUE)
head(recruits)

sort_recruits <- c("Number of Recruits", "Average Stars Per Recruit", "Average Rating Per Recruit")
sort_conference <- c("All FBS", "All Power 5", names(table(recruits$Conference)))
sort_conference
ui_recruits <- fluidPage(
  fluidRow(
    column(4, selectInput("year", label = "Recruiting Class Year: ", choices = 2012:2023)), 
    column(4, selectInput("recruit", label = "Sort By: ", choices = sort_recruits)),
    column(4, selectInput("conference", label = "Conference: ", choices = sort_conference))), 
  plotOutput("plot"),
  textOutput("model"),
  textOutput("pvalue"),
  dataTableOutput("table")
)

server_recruits <- function(input, output, session) {
  dataset <- reactive ({
    # sort by the selected conference
    if (input$conference == "All FBS") {
      data <- recruits
    } else if (input$conference == "All Power 5") {
      data <- recruits[recruits$Conference == "ACC" | recruits$Conference == "Big 12" | 
                       recruits$Conference == "Big Ten" | recruits$Conference == "Pac-12" | 
                       recruits$Conference == "SEC", ]
    } else {
      data <- recruits[recruits$Conference == input$conference, ]
    }
    # whether to show number of recruits, average stars per recruit, or average rating per recruit
    if (input$recruit == "Number of Recruits") {
      data <- data[, c("Team", paste("Recruits", input$year, sep = ""), "Wins2022")]
    } else if (input$recruit == "Average Stars Per Recruit") {
      data <- data[, c("Team", paste("AvgStars", input$year, sep = ""), "Wins2022")]
    } else {
      data <- data[, c("Team", paste("AvgRating", input$year, sep = ""), "Wins2022")]
    }
    # sort the data from greatest to lowest
    data[order(data[, 2], decreasing = TRUE), ]
  })
  
  linear_model <- reactive({
    lm(dataset()[, "Wins2022"] ~ dataset()[, 2])
  })
  
  output$plot <- renderPlot({
     plot(x = dataset()[, 2], y = dataset()[, 3], pch = 19, ylab = "Wins in 2022", 
          xlab = paste(input$recruit, "in", input$year), 
          main = paste("Wins in 2022 vs.", paste(input$recruit, "in", input$year), 
                       "for", input$conference, "Schools"))
    abline(lsfit(x = dataset()[, 2], y = dataset()[, 3]), col = "red")
  })
  
  output$model <- renderPrint({
    noquote(paste0("Line of best fit: (Wins in 2022) = ", linear_model()[[1]][2], " * (",
                   input$recruit, ") + ", linear_model()[[1]][1], \t))
  })
  
  output$pvalue <- renderPrint({
    p_value <- summary(linear_model())$coefficients[2, 4]
    if (p_value >= 0.05) {
      noquote(paste0(\t, "p-value = ", p_value, ". There is no significant relationship between", 
                     " a team's number of wins in 2022 and ", input$recruit, " in ",
                     input$conference))
    } else {
      noquote(paste0(\t, "p-value = ", p_value, ". There is a significant relationship between",
                     " a team's number of wins in 2022 and ", input$recruit, " in ",
                     input$conference, "."))
    }
  })
  
  output$table <- renderDataTable({
    data.frame("Rank" = 1:nrow(dataset()), dataset()[, c(1, 2)])
  })
}

recruits_app <- shinyApp(ui_recruits, server_recruits)

recruits_app
```

App that allows users to see transfer class rankings
```{r}
transfers <- read.csv("team transfers.csv", header = TRUE)
head(transfers)

sort_transfers <- c("Number of Transfers", "Average Stars Per Transfer")
sort_conference2 <- c("All FBS", "All Power 5", names(table(transfers$Conference)))
ui_transfers <- fluidPage(
  selectInput("year", label = "Transfer Class Year: ", choices = 2021:2023),
  selectInput("transfer", label = "Sort By: ", choices = sort_transfers),
  selectInput("conference", label = "Conference: ", choices = sort_conference2),
  dataTableOutput("table"), 
  plotOutput("plot")
)

server_transfers <- function(input, output, session) {
  dataset <- reactive({
    # choose what conference to show
    if (input$conference == "All FBS") {
      data <- transfers
    } else if (input$conference == "All Power 5") {
      data <- transfers[transfers$Conference == "ACC" | transfers$Conference == "Big 12" | 
                       transfers$Conference == "Big Ten" | transfers$Conference == "Pac-12" | 
                       transfers$Conference == "SEC", ]
    } else {
      data <- transfers[transfers$Conference == input$conference, ]
    }
    # whether to show number of transfers or average stars per transfer and for what year
    if (input$transfer == "Number of Transfers") {
      data <- data[, c("Team", paste("Transfers", input$year, sep = ""), "Wins2022")]
    } else {
      data <- data[, c("Team", paste("AvgStars", input$year, sep = ""), "Wins2022")]
    }
    # sort the data from greatest to lowest
    data[order(data[, 2], decreasing = TRUE), ]
  })
  output$table <- renderDataTable({
    data.frame("Rank" = 1:nrow(dataset()), dataset()[, c(1, 2)])
  })
  
  output$plot <- renderPlot({
     plot(x = dataset()[, 2], y = dataset()[, 3], pch = 19, ylab = "Wins in 2022", 
          xlab = paste(input$transfer, "in", input$year), 
          main = paste("Scatterplot of Wins in 2022 vs.", paste(input$transfer, "in", input$year), 
                       "for", input$conference, "Schools"))
    abline(lsfit(x = dataset()[, 2], y = dataset()[, 3]), col = "red")
  })
}

transfer_app <- shinyApp(ui_transfers, server_transfers)

transfer_app
```
















